## Basic Simulation with Same Parameters ##

### Problem ###

We start off with basic simulation with Rhipe. Our goal is to generate $N(0,1)$ variables. Suppose we want to generate $N$ random numbers and $N$ is very large.

### Simple R version ###

The simple R code for simulation is 

```{r eval=FALSE, tidy=FALSE}
N<-1000.
x<-rnorm(N)

```
The output will look like:( we are showing first 6 elements )
```
> head(x)
[1]  0.32987901 -0.28096045 -0.35665524  1.50459348 -1.77322751  0.02927208
> 
```

Also you will see this error message if $N$ is very large.
```
Error in rnorm(N) : invalid arguments
In addition: Warning message:
In rnorm(N) : NAs introduced by coercion

```

### RHIPE version ###

The entire code of RHIPE for simulation is:

```{r eval=FALSE, tidy=FALSE}
account <- "chakrav0"
N <- 2^20
R <- 2^10

##Generating data
map1 <- expression({
  lapply(seq_along(map.keys), function(r){
  	n <- N/R 
		value <- rnorm(n)
		rhcollect(r, value)
		rm(value)
	})
})
reduce1 <- expression(
  pre = {
    value <- numeric(0)
  },
  reduce = {
    value = c(value,unlist(reduce.values))
  },
  post = {
    rhcollect(1, value)
  }
)
job1 <- rhwatch(
	map = map1,
  reduce = reduce1,
	input = c(R, 20),
	output = rhfmt(file.path("/ln", account, "sim2", "data"), type="sequence"),
	jobname = "Generating data",
	readback = FALSE,
	parameters = list(N = N, R = R),
	noeval = TRUE
)
ex = rhex(job1, async=FALSE)

##read result from HDFS
rst <- rhread(file.path("/ln", account, "sim2", "data"))
rst
```

The mapreduce job in RHIPE is consist of a `map` expression, an optional `reduce` expression, and a execution funtion 
`rhwatch()`.

Let us look at the code carefully. First we decare the number of random variables $N$ we are going to simulate and the number of nodes $R$ in which we are going divide the simulation. We declare these variables outside any map or reduce expression.

Now let us look at the map expression. The map keys associated with this map expression are $1$ to $R$. For each of these keys we simulate $n=N/R$ number of observations. These $n$ simulations are values associated with the keys. So we have $R$ key-value pairs for the map expression.

Actually we can run the sumulation with just this map expression,

```{r eval=FALSE, tidy=FALSE}
account <- "chakrav0"
N <- 100
R <- 10

##Generating data
map1 <- expression({
  lapply(seq_along(map.keys), function(r){
  	n <- N/R 
		value <- rnorm(n)
		rhcollect(r, value)
		rm(value)
	})
})
job1 <- rhwatch(
	map = map1,
  input = c(R, 20),
	output = rhfmt(file.path("/ln", account, "sim2", "data"), type="sequence"),
	jobname = "Generating data",
	readback = FALSE,
	parameters = list(N = N, R = R),
	noeval = TRUE
)
ex = rhex(job1, async=FALSE)

##read result from HDFS
rst <- rhread(file.path("/ln", account, "sim2", "data"))
rst
```

The output of this code will be a collection of $R$ lists, each list will have a key
