## Manipulating Text Data ##

As mentioned previously, structured data stored in text files are ubiquitous nowadays.
RHIPE provides a powerful tool to manipulate these massive text files, and in this section we demonstrate how to use RHIPE to perform two importmant types of operations, namely, *subsetting* and *transforming* the data.

*
For instance, one might wish to *subset* a text file containing the information of both Southwest Airlines and Delta Airways and create two seperate text files instead, with one comprising the information of Southwest Airlines only and the other of Delta Airways only.

*
For instance, one might wish to transform the original text file by deleting certain columns or replacing the Airport Codes in the data to their full names instead.

For consistency, our discussion in this section continues to build on the Airline Dataset which we we have been utlizing in the preceeding sections so far.

### Subsetting Data (Under Construction) ###

Recall that the airline data take the form
```
BUGBUG
```
The carrier name is column 9. The carrier code for Southwest Airlines is WN, and the code for Delta Airways is DL. Only those rows with column 9 equal to WN or DL will be saved.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  tkn <- strsplit(unlist(map.values), ",")
  text <- do.call("rbind", tkn)
  text <- text[text[, 9] %in% c("WN", "DL"), , drop = FALSE]
  if (nrow(text) > 0)
    apply(text, 1, function(r) rhcollect(r[9], r))
})
```

Note that ```rhcollect()``` requires both a key and value.
However, since the key is not used, ```NULL``` is passed to the key argument and ```mapred.textoutputformat.usekey``` is set to ```FALSE``` so that the key is not written to disk.
By default RHIPE includes strings in quotes, but since we do not wish to do so here, we set ```rhipe_string_quote``` to ```''``` and ```mapred.field.separator``` to ```","``` as the original data is comma separated.
A partitioner is used to send all the flight information belonging to Southwest Airlines to one file and that belonging to Delta Airways to another.

```{r eval=FALSE, tidy=FALSE}
z <- rhwatch(
  map = map,
  reduce = rhoptions()$templates$identity,
  input = "tmp/airline/data/part-m-00005",
  output = "tmp/airline/southdelta",
  , inout = c("text", "text"),
  orderby = "char",
  part = list(lims = 1, type = "string"),
  mapred = list(
    mapred.reduce.tasks = 2,
    rhipe_string_quote = '',
    mapred.field.separator = ",",
    mapred.textoutputformat.usekey = FALSE)
  )
rhex(z)
```

The first output file contains
```
BUGBUG
```
and the sceond output file contains
```
BUGBUG
```

### Transforming Data (Under Construction) ###

In this example, we convert each airport code to their name equivalent. Airport codes can be found at the [JSM website](http://stat-computing.org/dataexpo/2009/the-data.html). When working with massive data, repeatedly used operations need to be as fast as possible. Thus we will save the airport code to airport name as a hash table using the ```new.env()``` function. Airport codes (origin and destination) are in columns 17 and 18. The setup expression loads this data set and creates a function that does the mapping.

```{r eval=FALSE, tidy=FALSE}
airport <- read.table("~/tmp/airports.csv", sep=",", header = TRUE, stringsAsFactors = FALSE)
aton <- new.env()
for(i in 1:nrow(airport)){
  aton[[ airport[i, "iata"] ]] <- list(ap = airport[i, "airport"], latlong = airport[i, c("lat", "long")])
}
rhsave(aton, file = "/tmp/airports.Rdata")

setup <- expression({
  load("airports.Rdata")
  co <- function(N){
    sapply(text[, N], function(r){
      o <- aton[[ r[1] ]]$ap
      if(is.null(o)) NA else sprintf('"%s"', o)
    })
  }
})
```

The map function will use the `aton` dictionary to get the complete names.
Note that removing the `sprintf()` function statement makes it much faster.

```{r eval=FALSE, tidy=FALSE}
map <- expression({
  tkn <- strsplit(unlist(map.values), ",")
  text <- do.call("rbind", tkn)
  text[, 17] <- co(17)
  text[, 18] <- co(18)
  apply(text, 1, function(r) { rhcollect(NULL, r) })
})

z <- rhwatch(map = map,ifolder = "/airline/data/2005.csv",
  ofolder = "/airline/transform", 
  , inout = c("text","text"),
  shared = c("/airport/airports.Rdata"),
  setup = setup,
  mapred = list(
    mapred.reduce.tasks = 0,
    rhipe_string_quote = '',
    mapred.field.separator = ",",
    mapred.textoutputformat.usekey = FALSE))
rhex(z)
```

The output is shown as follows.

```
BUGBUG
```